\plaintitle

### 2017-01-11 drafts/labor/discussion.tex
michael readthrough

\subsection{The Hazards of Predicting the Future}\label{sec:perilousProblemsPredicting}
The past isn't a perfect predictor for the future;
as \citeauthor{scholz2012digital} cautions,
``it would be wrong to conclude that in the realm of digital labor there is nothing new under the sun''~\cite{scholz2012digital}.
Our analysis is limited by the differences, foreseen and unforeseen, between historical piecework and modern on--demand work.
For example, unlike physical work environments, people can (and often do) make one--off contributions to online communities~\cite{mcinnis2016one}.
While we have attempted to identify all of the likely parallels and divergences between piecework and on--demand work,
we can't claim to have accounted for everything.

@@

\topic{On the other hand, while piecework's nascent years were grim, 
they precipitated a century of some of the most potent labor advocacy organizations in modern history~\cite{hart2013rise,mccallum2013global}.}
Even today, the \textit{geist} of the labor union revolution
inspires collective action and worker empowerment around the world:
Recently, in India, workers across the nation engaged in
the largest labor strike in human history
\cite{indiaStrikeRealNews}.
If labor advocacy groups can find ways to effect change in on--demand work as some have called for
\cite{futureUnions},
then the future of on--demand labor may follow
the same trajectory of worker empowerment that piecework saw.

>>>>

\subsection{The Hazards of Predicting the Future}\label{sec:perilousProblemsPredicting}
The past isn't a perfect predictor for the future;
as \citeauthor{scholz2012digital} cautions,
``it would be wrong to conclude that in the realm of digital labor there is nothing new under the sun''~\cite{scholz2012digital}.
Our analysis is limited by the differences, foreseen and unforeseen, between historical piecework and modern on--demand work.
For example, unlike physical work environments, people can (and often do) make one--off contributions to online communities~\cite{mcinnis2016one}.
While we have attempted to identify some likely parallels and divergences between piecework and on--demand work,
we can't claim to have accounted for everything.

@@

\topic{On the other hand, while piecework's nascent years were grim, 
they precipitated a century of some of the most potent labor advocacy organizations in modern history~\cite{hart2013rise,mccallum2013global}.}
Even today, the \textit{geist} of the labor union revolution
inspires collective action and worker empowerment around the world.
Recently, in India, workers across the nation engaged in
the largest labor strike in human history
\cite{indiaStrikeRealNews}.
If labor advocacy groups can find ways to effect change in on--demand work as some have called for
\cite{futureUnions},
then the future of on--demand labor may follow
the same trajectory of worker empowerment that piecework saw.

### 2017-01-11 drafts/labor/introduction.tex
michael readthrough

\topic{The past decade has seen a flourishing of computationally--mediated labor.}
The framing of work into modular components
has enabled computational hiring and management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed workers engage in work whenever their schedules allow,
often with little to no awareness of the broader context of the work, and
often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.

For years, such labor was limited to information work such as data annotation and surveys~\cite{CrowdsourcingUserStudies,movieSummarizationWu,yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
However, physically embodied work such as driving and cleaning have now spawned multiple online labor markets as well~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
In this paper we will use the term \textit{on--demand labor}, as \citeauthor to capture this pair of related phenomena:
first, \textit{crowd work}~\cite{crowdworkFuture}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
and second, \textit{gig work}~\cite{friedman2014workers}, often as platforms for one--off jobs, like driving, courier services, and administrative support.
\ali{I'm a little worried about people offering contradicting terms, but the argument I \textit{want} to make is that
these terms (``gig work'' and ``crowd work'') were sort of born out of marketing and business interests (just like ``sharing economy''),
and as a result they're somewhat fuzzy and imprecise (for example, see~\cite{de2015crowdsourcing}); what we're trying to do is
place these terms where they're most accurate and tightly scoped, so that their meaningfulness is retained;
then we use  ``on--demand work'' to capture this work more holistically,
much more carefully selecting the terminology for academic discourse.

@@

The realization that complex goals can be accomplished by directing crowds of workers has spurred industry to flock to sites of labor
such as AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems that enable new forms of production
(e.g. \cite{bernsteinSoylent,vizwiz,paolacci2010running}) and pursuing social scientific inquiry into the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
This research has called out identified the sociality of gig work~\cite{crowdcollab},
as well as the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
Others have focused on the \textit{outcomes} of this frustration,
reflecting on the resistance that workers express against digitally--mediated labor markets~\cite{uberAlgorithm,dynamo}.

@@

    {Workers} &
    \small{Firms antagonized and exploited workers, leading workers to support one another independently, ultimately resulting in strong advocacy groups counterbalancing firms.} &
    \small{The features of piecework (independence and transience) were both the fulcrum managers used to exploit workers as well as the focal point around which workers bonded.} &
    \small{While worker frustrations are similar, the decentralized nature of on--demand work will limit collective action until there exist platforms to coordinate and exert pressure.} \\ 
    \bottomrule
  \end{tabularx}
  \caption{Piecework and on--demand work have both wrestled with questions of how complex work can get, how finely--sliced tasks can become, and what the workplace looks like for workers. We connect piecework's history (left) to the mechanisms that impacted it (center) to derive predictions for modern on--demand work (right). }
\end{table*}

\subsection{Piecework as a lens to understand crowdsourcing}
In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework},
a work and payment structure which breaks tasks down into discrete jobs,
wherein payment is made for \textit{output}, rather than for \textit{time}.
Piecework use as a lens on on--demand work is not new.
In \citeyear{crowdworkFuture}, for example,
\citeauthor{crowdworkFuture} referenced crowd work as piecework briefly
as a loose analogy~\cite{crowdworkFuture}.
But more than this,
the framing of on--demand labor as a reinstantiation of piecework % google said reinstantion is a word
gives us years of historical material to make sense of this new form of work, and
allows us to reflect on--demand work through a theoretical lens, informed by years of rigorous, empirical research.

More concretely, by looking at on--demand labor as
an instantiation (or even a continuation) of piecework,
and by interrogating patterns that the historical literature identifies, we can achieve three goals.
First, we can make sense of past events as part of a much larger series of an interrelated phenomena.
Second, we can reflect on differences in the features that impacted piecework historically and on--demand work today.
And third, to some extent, we can use these differences to offer some predictions of what on--demand labor researchers,
and the workers,
might expect to see on the horizon.
For example, we will draw on the piecework literature regarding task decomposition,
which was historically limited by shortcomings in measurement and instrumentation, and
leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
--- namely, by enabling precise tracking and measurement via algorithms and software.

We organize this paper as follows:
first, we review the definition and history of piecework
to make clear the analogy to on--demand work;
and second, we interrogate the three major research questions above using the lens of piecework. 
For each question, we will
contrast the perspective the piecework scholarship offers with on--demand labor's body of research,
identify similarities and differences, and then
offer predictions for on--demand work.

>>>>

\topic{The past decade has seen a flourishing of computationally--mediated labor.}
A framing of work into modular, pre-defined components
enables computational hiring and management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
In this regime, distributed workers engage in work whenever their schedules allow,
often with little to no awareness of the broader context of the work, and
often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.

For years, such labor was limited to information work such as data annotation and surveys~\cite{CrowdsourcingUserStudies,movieSummarizationWu,yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
However, physically embodied work such as driving and cleaning have now spawned multiple online labor markets as well~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
In this paper we will use the term \textit{on--demand labor}, to capture this pair of related phenomena:
first, \textit{crowd work}~\cite{crowdworkFuture}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
and second, \textit{gig work}~\cite{friedman2014workers}, often as platforms for one--off jobs, like driving, courier services, and administrative support.
\ali{I'm a little worried about people offering contradicting terms, but the argument I \textit{want} to make is that
these terms (``gig work'' and ``crowd work'') were sort of born out of marketing and business interests (just like ``sharing economy''),
and as a result they're somewhat fuzzy and imprecise (for example, see~\cite{de2015crowdsourcing}); what we're trying to do is
place these terms where they're most accurate and tightly scoped, so that their meaningfulness is retained;
then we use  ``on--demand work'' to capture this work more holistically,
much more carefully selecting the terminology for academic discourse.

@@

The realization that complex goals can be accomplished by directing crowds of workers has spurred industry to flock to sites of labor
such as AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems that enable new forms of production
(e.g. \cite{bernsteinSoylent,vizwiz,paolacci2010running}) and pursuing social scientific inquiry into the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
This research has identified the sociality of gig work~\cite{crowdcollab},
as well as the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
Others have focused on the \textit{outcomes} of this frustration,
reflecting on the resistance that workers express against digitally--mediated labor markets~\cite{uberAlgorithm,dynamo}.

@@

    {Workers} &
    \small{Firms antagonized and exploited workers, leading workers to support one another independently, ultimately resulting in strong advocacy groups counterbalancing firms.} &
    \small{The features of piecework (independence and transience) were both the fulcrum managers used to exploit workers as well as the focal point around which workers bonded.} &
    \small{While worker frustrations are similar, the decentralized nature of on--demand work will limit collective action until there exist platforms to coordinate and exert pressure.} \\ 
    \bottomrule
  \end{tabularx}
  \label{tab:overview}
  \caption{Piecework and on--demand work have both wrestled with questions of how complex work can get, how finely--sliced tasks can become, and what the workplace will look like for workers. We connect piecework's history (left) to the mechanisms that determined its outcomes to these three questions (center) in order to derive predictions for modern on--demand work (right). }
\end{table*}

\subsection{Piecework as a lens to understand on--demand work}
In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework},
a work and payment structure which breaks tasks down into discrete jobs,
wherein payment is made for \textit{output}, rather than for \textit{time}.
We are not the first to relate on--demand work to piecework: in \citeyear{crowdworkFuture}, for example,
\citeauthor{crowdworkFuture} referenced crowd work as piecework briefly
as a loose analogy~\cite{crowdworkFuture}.
Our goal in this paper is to inspect the relationship much more closely.
But more than this,
the framing of on--demand labor as a reinstantiation of piecework 
gives us years of historical material to help us make sense of this new form of work, and
allows us to reflect on--demand work through a theoretical lens that is informed by years of rigorous, empirical research.

More concretely, by interrogating on--demand labor as
an instantiation or even a continuation of piecework, we can make sense of past events as part of a much larger series of an interrelated phenomena (Table~\ref{tab:overview}).
We can reflect on differences in the features that impacted piecework historically and on--demand work today.
And, to some extent, we can use these differences to offer some predictions of what on--demand work researchers
and workers themselves
might expect to see on the horizon.
For example, we will draw on the piecework literature regarding task decomposition,
which was historically limited by shortcomings in measurement and instrumentation, and
leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
--- namely, by enabling precise tracking and measurement via algorithms and software.

We organize this paper as follows:
first, we review the definition and history of piecework
to make clear the analogy to on--demand work;
and second, we examine the three major research questions above using the lens of piecework. 
For each question, we will
contrast the perspective the piecework scholarship offers with on--demand labor's body of research,
identify similarities and differences, and then
offer predictions for on--demand work.

### 2017-01-11 drafts/labor/piecework_lit.tex
michael readthrough

\topic{Piecework's history traces back further perhaps than most would expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
to calculate equations in order
to better--predict the trajectories of various celestial bodies in the night sky~\cite{grier2013computers}.
George Airy was perhaps the first to rigorously put piecework--style decomposition to work;
by breaking complex calculations into constituent parts, and
training young men to solve simple algebraic problems,
Airy could distribute work to many more people than could otherwise complete the full calculations.

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took hold in farm work~\cite{hughRaynbirdTaskWork}, % hold foothold
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491}.
At the turn of the \nth{20} century, \citeauthor{taylor1914principles}
--- a mechanical engineer with a fascination in work efficiency ---
began studying and formalizing the decomposition, tracking, and management of tasks.
In \citeyear{taylor1914principles} he published \citetitle{taylor1914principles},
concretizing an idea that had nebulously been forming,
and which he had been working out himself, for years~\cite{taylor1914principles}.
Scientific Management (and Fordism) thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.

@@

\topic{By \citeyear{hughRaynbirdTaskWork} we find
a concise definition of piecework
in \citeauthor{hughRaynbirdTaskWork}'s essay on piecework,
particularly driven toward encapsulating the manual labor of farmwork.}
He does this by contrasting two paradigms ---
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} defines it through examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing offers an intuitive sense of piecework;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

@@

\topic{We see this sense of independence in myriad times, locales, and industries.}
\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls''
--- teenage women who assembled matchsticks in the late \nth{19} century in London. % added that these women were in London since Chinmay critiqued that we're glossing over the labor advocacy movement that grew in Europe, which revealed that we completely left out the fact that this all happened in England.
\ali{Chinmay: but still not convinced; this is a post hoc ergo propter hoc argument}
Of interest was their reputation ``\dots~for generosity, independence, and protectiveness,
but also for brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
\citeauthor{10.2307/27508091} document piecework from 1850 through 1930 in Australia,
finding similar notions of independence and autonomy among piecework newspaper compositors:
``If a piece--work compositor~\dots~decided that he did not want to work on a particular day or night, the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
This sense of independence and autonomy appears to be a common thread of piecework.

@@

\topic{Piecework opened the door for people who previously couldn't participate in the labor market to do so, and to acquire job skills incrementally.}
During World War II, women received training in narrow subsets of more comprehensive jobs, enabling work in capacities similar to conventional (male) workers~\cite{hart2013rise}.
Women previously had virtually no opportunities
to engage in engineering and metalworking apprenticeships as men did;
now, they
could be trained quickly on narrowly scoped tasks,
demonstrate proficiency, and become experts.

@@

Viewing on--demand work as a modern instantiation of piecework is relatively straightforward by this definition.
First, platforms such as Amazon Mechanical Turk (AMT), Uber, Upwork, and TaskRabbit pay by the task, though some mix systems in similar ways to the Rowan system's combination of piece rate and time rate pay.
Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well.

@@

\topic{Piecework found its way into the war effort during World War~II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a mostly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
and correspondingly assigned to that or a similar task.
``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
would have been a pieceworker~\cite{davies2014origins}. \ali{was? would have been?}

>>>>

\topic{Piecework's history traces back further perhaps than most would expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
to calculate equations in order
to better--predict the trajectories of various celestial bodies in the night sky~\cite{grier2013computers}.
In the first half of the \nth{19} century, George Airy was perhaps the first to rigorously put piecework--style decomposition to work;
by breaking complex calculations into constituent parts, and
training young men to solve simple algebraic problems,
Airy could distribute work to many more people than could otherwise complete the full calculations.

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took hold in farm work~\cite{hughRaynbirdTaskWork}, % hold foothold
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491} by the mid-19th century.
\topic{By \citeyear{hughRaynbirdTaskWork} we find
a concise definition of piecework
in \citeauthor{hughRaynbirdTaskWork}'s essay on piecework,
particularly driven toward encapsulating the manual labor of farmwork.}
He does this by contrasting two paradigms ---
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} defines it through examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing offers an intuitive sense of piecework;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

At the turn of the \nth{20} century, \citeauthor{taylor1914principles}
--- a mechanical engineer with a fascination in work efficiency ---
began studying and formalizing the decomposition, tracking, and management of tasks.
In \citeyear{taylor1914principles} he published \citetitle{taylor1914principles},
concretizing an idea that had nebulously been forming,
and which he had been working out himself, for years~\cite{taylor1914principles}.
Scientific Management (and Fordism) thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.

@@

@@

\topic{We see this sense of independence in myriad times, locales, and industries.}
\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls'', teenage women who assembled matchsticks in the late \nth{19} century in London. % added that these women were in London since Chinmay critiqued that we're glossing over the labor advocacy movement that grew in Europe, which revealed that we completely left out the fact that this all happened in England.
\ali{Chinmay: but still not convinced; this is a post hoc ergo propter hoc argument}
Of interest was their reputation ``\dots~for generosity, independence, and protectiveness,
but also for brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
\citeauthor{10.2307/27508091} document piecework from 1850 through 1930 in Australia,
finding similar notions of independence and autonomy among piecework newspaper compositors:
``If a piece--work compositor~\dots~decided that he did not want to work on a particular day or night, the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
This sense of independence and autonomy appears to be a common thread of piecework.

@@

\topic{Piecework opened the door for people who previously couldn't participate in the labor market to do so, and to acquire job skills incrementally.}
During World War II, women received training in narrow subsets of more comprehensive jobs, enabling work in capacities similar to conventional (male) workers~\cite{hart2013rise}.
Women previously had virtually no opportunities
to engage in engineering and metalworking apprenticeships as men did;
now, they
could be trained quickly on narrowly scoped tasks,
demonstrate proficiency, and become experts.
``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
would have been a pieceworker~\cite{davies2014origins}.

@@

Viewing on--demand work as a modern instantiation of piecework is relatively straightforward by this definition.
First, platforms such as Amazon Mechanical Turk (AMT), Uber, Upwork, and TaskRabbit pay by the task, though some mix systems in similar ways to the Rowan system's combination of piece rate and time rate pay.
Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
Given this alignment, many of the same historical properties of piecework will apply to on--demand work as well.

@@

\topic{Piecework found its way into the war effort during World War~II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a mostly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
and correspondingly assigned to that or a similar task.
